Supplemental Information B: Full Annotated Analysis and Reproducible Code

### 2021-10-10 Supplemental-Material-B-Annotated-Analysis.Rmd
dtWorker munge added

@@

## Data Exclusions

>>>>

  df <- df %>%
    mutate(
      PID = as.numeric(factor(ExternalReference)), # participant ID
      TID = measureID-1, # time ID with t0 = 0 for meaningfull intercept interpretations
      date = substr(StartDate,1,10), # awkward way of extracting date (best converted to )
      time = substr(StartDate,12,19), # awkward way of extracting time
      daynum = as.numeric(factor(date)), # all days as numeric for ordering
      daycor = ifelse(daytime=="morning" & period_to_seconds(hms(time))<period_to_seconds(hms("12:00:00")) | daytime=="afternoon" & period_to_seconds(hms(time))<period_to_seconds(hms("19:00:00")),daynum-1,daynum), # correctly identify which date the questionnaire is about
      daycor.lead = sprintf("%02d", daycor),
      daytime.lt = ifelse(daytime=="morning","a","b"), # morning / afternoon to a / b 
      day_time = paste(daycor.lead, daytime.lt, sep="_"), # combine day id with morning / afternoon
      session = as.numeric(factor(day_time)), # day and time identifier as numeric id
      SubTime = chron::times(time.0),
      time.daily = as.character(time.daily),
      PPDate = as.Date(df$date.daily),
      number = replace_na(number, 0),
      NonDutchNum = replace_na(NonDutchNum, 0)
    )

@@

## Calculate needed transformations

```{r newVarsWorkers}
df <- dtWorker$clean

# Time and Date Variables
  # remove seconds from afternoon time
  df$SubTime[df$daytime == "afternoon"] = paste0(substring(as.character(df$time.0[df$daytime == "afternoon"]),4,8),":00") 
  df$time.daily[df$daytime == "afternoon" & !is.na(df$time.daily!="<NA>")] = paste0(substring(as.character(df$time.daily[df$daytime == "afternoon" & !is.na(df$time.daily!="<NA>")]),4,8),":00")

  # Correct morning / afternoon date where survey was collected the day after to indicate the correct date that was targeted
  df$PPDate[df$SubTime < "11:50:00" & df$daytime == "morning"] = df$PPDate[df$SubTime < "11:50:00" & df$daytime == "morning"]-1
  df$PPDate[df$SubTime < "18:50:00" & df$daytime == "afternoon"] = df$PPDate[df$SubTime < "18:50:00" & df$daytime == "afternoon"]-1

# Mood scales
  df$calmness.daily = rowSums(df[, c("calmness1", "calmness2")], na.rm = T)
  df$alertness.daily = rowSums(df[, c("alertness1", "alertness2")], na.rm =T)
  df$valence.daily = rowSums(df[, c("valence1", "valence2")], na.rm = T)

# Need scales
  df$keyMotiveFulfilled = rowSums(df[,c("keymotive_fulfillemt_1","keyMotive_noInt_fulf_1")], na.rm=T)
  df$autonomy.daily.all = rowSums(df[,c("autonomy_1","autonomy_NoInt_1")], na.rm=T)
  df$competence.daily.all = rowSums(df[,c("competence_1","competence_NoInt_1")], na.rm=T)
  #cor(df$relatedness_other_1, df$relatedness_self_1,use="complete.obs")
  df$relatedness.daily.all = rowMeans(df[,c("relatedness_other_1","relatedness_self_1","relatedness_1_NoInt_1")], na.rm=T)
  df$relatedness.daily.int = rowMeans(df[,c("relatedness_other_1","relatedness_self_1")], na.rm=T)

# summarize by participant (check that everything is within pp might not be the case for )
  between <- df %>%
    group_by(ExternalReference) %>%
    mutate(
      CtContactNL = sum(Contact_dum),
      CtContactNonNl = sum(inNonDutch),
      CtContactNLAll = sum(number),
      CtContactNonNlAll = sum(NonDutchNum),
      AvKeyNeed = mean(keyMotiveFulfilled, na.rm=T),
      AvKeyNeedInt = mean(keymotive_fulfillemt_1, na.rm=T),
      AvKeyNeedNoInt = mean(keyMotive_noInt_fulf_1, na.rm=T),
      AvAutonomy = mean(autonomy.daily.all, na.rm=T),
      AvCompetence = mean(competence.daily.all, na.rm=T),
      AvRelatedness = mean(relatedness.daily.all, na.rm=T),
      AvThermo = mean(thermometerDutch_1, na.rm=T),
      AvWB = mean(ExWB_1, na.rm=T)) %>%
    ungroup() %>%
    mutate(
      CtContactNL_c = scale(CtContactNL, scale = FALSE),
      AvKeyNeedInt_c = scale(AvKeyNeedInt, scale = FALSE),
      AvKeyNeed_c = scale(AvKeyNeed, scale = FALSE),
      CtContactNL_z = scale(CtContactNL, scale = TRUE),
      AvKeyNeedInt_z = scale(AvKeyNeedInt, scale = TRUE),
      AvKeyNeed_z = scale(AvKeyNeed, scale = TRUE)
    )

  warning("some variable transformations (esp. _c and _z) might be across all participants (i.e., not within PP).")

  dtWorker$munge <- between

  rm(df, between)

  #save(df.btw, file = "data/processed/df.btw.RData")  
  #write_sav(df.btw, "data/processed/MT_clean-merged_pre-post.sav")

# export data to Mplus
  # df.mplus = remove_all_labels(select(df, PID, session, thermometerDutch_1, inNonDutch, Contact_dum, keyMotiveFulfilled, autonomy.daily.all, competence.daily.all, relatedness.daily.all))
  # names(df.mplus)= c("PID", "session", "att", "intin", "intout", "keymot", "aut", "comp", "rel")
  # mplus = df.mplus[order(df.mplus$PID, df.mplus$session),]
  # mplus.intcont = mplus[mplus$intout==1,]
  # prepareMplusData(mplus.intcont, "data/processed/dynamic-subset-intonly.dat")

```{r newVarsStudents}
df <-

# Add ID variables
  df$PID = as.numeric(factor(df$session)) # participant ID

# order time
  # move this here? 
  # date.period.order <- unique(raw$dat.daily$date_period)
  date.period.order <- .GlobalEnv$date.period.order # <- @Max: Do you still need this? Order should be fine natively from FormR

  df$TID <- factor(df$date_period, levels = date.period.order)
  df$TIDnum <- as.numeric(df$TID) #get numeric TID

  # check whether time ordering worked
  df_ordered <- df %>%
    arrange(PID, TID) #%>%
    #View()

# Interaction as Factor
  df$interaction.f <- factor(df$Interaction, levels = c("no interaction", "Dutch", "Non-Dutch"))
  df$intNL <- ifelse(df$Interaction=="Dutch",1,0)
  df$intNonNL<- ifelse(df$Interaction=="Non-Dutch",1,0)

# ------------------------------------------------------------------------------------------------------------- 
#                                       Combine Variables
# -------------------------------------------------------------------------------------------------------------
source("lib/fun.panel.R")
# Relatedness
  .GlobalEnv$pairs.panels.new(df[c("RelatednessSelf","RelatednessOther")], 
                              labels = c("I shared information about myself.", "X shared information about themselves."))
  df$RelatednessInteraction <- rowMeans(df[c("RelatednessSelf","RelatednessOther")], na.rm=T)
  df$RelatednessInteraction[df$RelatednessInteraction == "NaN"] <- NA
# Relatedness Overall (JANNIS NOT SURE THESE ARE CORRECT, CHANGE ROWS?; J: Changed "NaN" in df$RelatednessInteraction to NA() should work now)
  df$Relatedness <- rowMeans(df[,c("RelatednessInteraction", "RelatednessNoInteraction")], na.rm=T)
# Pro-Sociality
  df$ProSo <- rowMeans(df[,c("ProSo1", "ProSo2", "ProSo3", "ProSo4")], na.rm=T)
# Anti-Sociality
  df$AntiSo <- rowMeans(df[,c("AntiSo1", "AntiSo2", "AntiSo3", "AntiSo4")], na.rm=T)

# ------------------------------------------------------------------------------------------------------------- 
#                                 Create Time Variables
# -------------------------------------------------------------------------------------------------------------
createTimeDf <- function(tmp) {
# make dataframe for all variables that need t-1
# @Max: LAGS AREN'T REALLY MEANINGFUL BECAUSE OF THE TIME INTERVAL DIFFERENCES, WE WILL DO THIS IN DSEM (MPLUS).
  tmp <- data.frame(KeyNeedFullfillment = tmp$KeyNeedFullfillment,
                    ProSo = tmp$ProSo,
                    AntiSo = tmp$AntiSo,
                    AttitudesDutch = tmp$AttitudesDutch)
    length <- ncol(tmp) # get the length of tmp for later renaming of columns

  # use dplyr to create time series
    tmp <- tmp %>% mutate_each(funs(lag(.,1))) %>% cbind(tmp, .)
    colnames(tmp)[(length+1):ncol(tmp)] <- paste(colnames(tmp)[1:length], "Lag1", sep = "") #rename after specific length (add one so new columns get renamed after that)
  tmp
}

# split df per participants and run function
  PP <- split(df, df$PID)
  PP <- lapply(PP, createTimeDf); rm(createTimeDf)

# add variables back to df
  tmp <- do.call(rbind.data.frame, PP)
    colnames(tmp) <- paste(colnames(tmp), "Calc", sep = "")
df <- cbind(tmp, df);rm(tmp)

# ------------------------------------------------------------------------------------------------------------- 
#                                 Add Variables related to interaction partner
# -------------------------------------------------------------------------------------------------------------

# create function for later lapply
createIntPartDf <- function(inp) {

  # prepare the dataframe so that we can forloop over it later
    tmp <- data.frame(CC = as.character(inp$CC),
                    NewCC = as.character(inp$NewCC),
                    NewName = as.character(inp$NewName),
                    NewCloseness = inp$NewCloseness,
                    NewGender = inp$NewGender,
                    NewEthnicity = as.character(inp$NewEthnicity),
                    NewRelationship = as.character(inp$NewRelationship))

    tmp$CC2 <- recode(tmp$CC, "'SOMEONE ELSE' = NA")
    tmp$CC2 <- ifelse(tmp$CC == 1 | tmp$CC == "SOMEONE ELSE", as.character(tmp$NewName), as.character(tmp$CC2)) 
    # maybe add [[:space:]]\b to remove space before word boundary or ^[[:space:]] to remove space in the beginning of a string
    tmp$CC2 <- gsub("^[[:space:]]", "", tmp$CC2)
    tmp$NewName <- gsub("^[[:space:]]", "", tmp$NewName)

  # open the variables that will be filled up in the foor-loop
    tmp$closeness <- rep(NA, nrow(tmp))
    tmp$gender <- rep(NA, nrow(tmp))
    tmp$ethnicity <- rep(NA, nrow(tmp))
    tmp$relationship <- rep(NA, nrow(tmp))

    # Run the for-loop. It finds the variables related to the name of the interaction partner. If there is a repeating interaction
    # partner (i.e. CC2) it takes the value (i.e. NewCloseness) from the first interaction (i.e. NewName)
      for (i in 1:nrow(tmp)) {
         if (is.na(tmp$CC2[i])) {
           next
         } else {
           tmp$closeness[i] <- na.omit(tmp$NewCloseness[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] #find closeness where CC2 matches NewName (na.omit + [1] to get the number)
           tmp$gender[i] <- na.omit(tmp$NewGender[as.character(tmp$CC2[i]) == as.character(tmp$NewName)])[1] #(na.omit + [1] to get the number and not the rest of the na.omit list)
           tmp$ethnicity[i] <- na.omit(as.character(tmp$NewEthnicity[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1] #PROBLEM IS THAT THERE ARE TOO MANY NA's: Difficult to deal with
           tmp$relationship[i] <- na.omit(as.character(tmp$NewRelationship[as.character(tmp$CC2[i]) == as.character(tmp$NewName)]))[1]
         }
      }

    out <- tmp
    out
}

# split df per participants and run function
  PP <- split(df, df$PID)
  PP <- lapply(PP, createIntPartDf); rm(createIntPartDf)

# add variables back to df
  remergePP <- do.call(rbind.data.frame, PP)
    colnames(remergePP) <- paste(colnames(remergePP), "_Calc", sep = "")
df <- cbind(remergePP, df);rm(remergePP)
print("Test whether newly calculate df and old df have the same order:")
print(table(as.character(df$CC_Calc) == as.character(df$CC)))
print("If we also add the NAs we get the exact number:")
print(sum(is.na(df$CC_Calc))+table(as.character(df$CC_Calc) == as.character(df$CC)))

# ------------------------------------------------------------------------------------------------------------- 
#                                 Center Relevant Variables
# -------------------------------------------------------------------------------------------------------------
  # center within participant
  # @Max: I removed the lagged variables so we can do it proberly in MPlus
    df$ProSo_c <- ave(df$ProSoCalc, df$PID, FUN=function(x) scale(x, center = T, scale = F))
    #df$ProSoLag1_c <- ave(df$ProSoLag1Calc, df$PID, FUN=function(x) scale(x, center = T, scale = F))
    df$AntiSo_c <- ave(df$AntiSoCalc, df$PID, FUN=function(x) scale(x, center = T, scale = F))
    #df$AntiSoLag1_c <- ave(df$AntiSoLag1Calc, df$PID, FUN=function(x) scale(x, center = T, scale = F))
    df$KeyNeedFullfillment_c <- ave(df$KeyNeedFullfillmentCalc, df$PID, FUN=function(x) scale(x, center = T, scale = F))
    #df$KeyNeedFullfillmentLag1_c <- ave(df$KeyNeedFullfillmentLag1Calc, df$PID, FUN=function(x) scale(x, center = T, scale = F))

    df$closeness_c <- ave(df$closeness_Calc, df$PID, FUN=function(x) scale(x, center = T, scale = F))

# variancec
  df$var.AntiSo <- ave(df$AntiSo, df$PID, FUN=var)

# send back to global environment
  .GlobalEnv$df.full = df
